---
title: "Time Series Forecasting"
output: html_notebook
---

# Goal

Predict the next 3 months of daily quantities of vehicle parts ordered. 

# 1 . Set Up

## 1.1 Load Libraries

```{r}
# Main Packages
library(tidyverse)
library(lubridate)
library(tidyquant)
library(plotly)
library(ggthemes)

# Time Based Feature Extraction
library(timetk)

# Holidays
library(timeDate)

# Weather Data
library(riem)

# Forecasting
library(forecast)
library(sweep)

# Machine Learning 
library(parsnip)
library(yardstick)
library(glmnet)
```

## 1.2 Load Data

```{r}
# Load data
sales_data_raw <- read_csv('sales_data_sample.csv') 
data <- sales_data_raw
data
```


# 2. Data Preprocessing

## 2.1 Parts Table

```{r}
# 
parts_d_tbl_all <- data %>%
  select(ORDERDATE, ORDERNUMBER, QUANTITYORDERED, PRICEEACH, 
         PRODUCTLINE, PRODUCTCODE, STATUS) %>%
  mutate(ORDERDATE = mdy_hm(ORDERDATE),
         ORDERDATE = as_datetime(ORDERDATE)) %>%
  group_by(ORDERDATE, PRODUCTCODE) %>%
  summarise(QUANTITYORDERED)

parts_d_tbl <- parts_d_tbl_all %>%
  group_by(ORDERDATE) %>%
  summarise(TOTAL = sum(QUANTITYORDERED))
```

## 2.2 Quantity Ordered Over Time

```{r}
# Plot it
g <- parts_d_tbl %>%
  ggplot(aes(ORDERDATE, TOTAL)) + 
  geom_line(alpha = 0.5, color = '#2c3e50') + 
  geom_smooth(method = "loess", span = 0.5) + 
  theme_tq() + 
  scale_color_tq() + 
  labs(x="",y="Quantity",title=" Quantity Ordered Over Time")

ggplotly(g) %>%
  layout(xaxis = list(rangeslider = list(type = "date")))
```

> Note the seasonality in the data. 

## 2.3 Data Augmentation

- Add time-based features using `timetk`
- Add holidays using `timedate`

```{r}
# Add time-based features
parts_d_tbl <- parts_d_tbl %>%
  tk_augment_timeseries_signature() %>%
  select(ORDERDATE, TOTAL, index.num, year, half, quarter, month.lbl, day, wday.lbl)

# Create holidays
holidays <- holidayNYSE(year = c(2003, 2004, 2005)) %>% ymd()

# Add holidays feature
parts_d_tbl <- parts_d_tbl %>%
  mutate(holiday = case_when(
    ORDERDATE %in% holidays ~ 1, 
    TRUE ~ 0
  ))
```

# 3. Exploratory Analysis

## 3.1 Autocorrelation


```{r}
autocorrelate <- function(data, value, lags = 0:20) {
    
    value_expr <- enquo(value)
    
    acf_values <- data %>%
        select(!! value_expr) %>%
        pull() %>%
        acf(lag.max = tail(lags, 1), plot = FALSE) %>%
        .$acf %>% # Remove text
        .[,,1]
    
    ret <- tibble(acf = acf_values) %>%
        rowid_to_column(var = "lag") %>% # Add lag columns
        mutate(lag = lag - 1) %>% 
        filter(lag %in% lags) 
    
    return(ret)
} 

g <- parts_d_tbl %>%
    autocorrelate(TOTAL, lags = 0:nrow(.)) %>%
    ggplot(aes(lag, acf)) +
    geom_point(alpha = 0.5, color = "#2c3e50") +
    expand_limits(y = c(-1, 1)) +
    theme_tq() +
    labs(title = "Autocorrelation")

ggplotly(g)
```

## 3. Modeling

## 3.1 Train-Test-Split

```{r}
train_test_split_date <- ymd("2005-03-01")

train_tbl <- parts_d_tbl %>%
    filter(ORDERDATE < train_test_split_date)

test_tbl <- parts_d_tbl %>%
    filter(ORDERDATE >= train_test_split_date)

# What Autocorrelation can we use in a multivariate model?
nrow(test_tbl)
```


## 3.2 Evaluate 60-Day Moving Average

```{r}
moving_average_train_tbl <- train_tbl %>%
    select(ORDERDATE, TOTAL) %>%
    mutate(moving_average = rollmean(TOTAL, k = 30, na.pad = TRUE, align = "right")) 

g <- moving_average_train_tbl %>%
    bind_rows(test_tbl %>% select(ORDERDATE, TOTAL)) %>%
    fill(moving_average, .direction = "down") %>%
    mutate(ORDERDATE = as.Date(as.character(ORDERDATE))) %>%
    ggplot(aes(ORDERDATE, TOTAL)) +
    geom_vline(xintercept = train_test_split_date, color = "red") +
    geom_point(color = "#2c3e50") +
    geom_line(aes(y = moving_average), size = 1, color = "blue") +
    theme_tq() 

ggplotly(g)

test_tbl %>%
    select(TOTAL) %>%
    mutate(moving_average = moving_average_train_tbl %>% 
               tail(1) %>% 
               pull(moving_average)) %>%
    mae(TOTAL, moving_average)
```





















